---
title: "Alzheimer's Proteomics Analyisis"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Differential Expression plots for Alzheimer's Proteomics experiments

This analysis consists of proteins pulldown from Tg P301s mice and WT controls.
There are also tdTtomato control animals which represents all background proteins.

```{r Loading libraries and setting directory, message=FALSE}
library(EnhancedVolcano)
library(MSstatsTMT)
library(UniProt.ws)
library(MSstats)
library(readr)

setwd("/gstore/scratch/u/lucast3/Alzheimers_Proteomics/")
```

## Tg P301s BioID vs All other control animals

Here we are making the comparison to see which proteins are enriched in AD animals over controls.

```{r tgBioId vs all, echo=TRUE, message=FALSE, warning=FALSE}
# Read in Differential expression data
data <- read_delim("Tg_BioIDvsAll.csv",
                   delim = ";", escape_double = FALSE,
                   col_types = cols(log2FC = col_number()),
                   locale = locale(decimal_mark = ",", grouping_mark = "."),
                   trim_ws = TRUE)
```

#### Including Plots

Volcano plot for enriched proteins

```{r tgBioId Volcano, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11.75, fig.height=10.04}
x <- data$GeneName
data <- data[-which(startsWith(as.character(x),"Krt")==TRUE),]
y1 <- ceiling(max(-log10(data$adj.pvalue))*2)/2


plot <- EnhancedVolcano(data,
                lab = data$GeneName,
                x = 'log2FC',
                y = 'adj.pvalue',
                title = 'TgBioID vs. All', caption='',
                subtitle = paste("FDR cutoff = 0.05 \nFC cutoff = 1"),
                pCutoff = 0.05,
                FCcutoff = 1,
                pointSize = 1.0,
                labSize = 4.0,
                col=c('black', 'black', 'black', 'red3'),
                ylim = c(0,y1),
                ylab = bquote(~-Log[10]~ 'FDR'),
                drawConnectors = TRUE,border = 'full',colConnectors = 'grey30',
                widthConnectors = .5, borderWidth = 1.0, borderColour = 'black') + theme(legend.position="none")

print(plot)

```

## WT BioID vs vs All other  animals

```{r WT BioId vs all, message=FALSE, warning=FALSE, echo=TRUE}
# Read in Differential expression
data <- read_delim("WtBvsALL.csv", 
                       delim = ";", escape_double = FALSE,col_types = cols(log2FC = col_number()),
                       locale = locale(decimal_mark = ",", grouping_mark = "."),
                       trim_ws = TRUE)
```

```{r Wt BioID vs all Volcano, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11.75, fig.height=10.04}
x <- data$GeneName
data <- data[-which(startsWith(as.character(x),"Krt")==TRUE),]
y1 <- ceiling(max(-log10(data$adj.pvalue))*2)/2
plot <- EnhancedVolcano(data,
                lab = data$GeneName,
                x = 'log2FC',
                y = 'adj.pvalue',
                title = 'WtBioID vs. All', caption='',
                subtitle = paste("FDR cutoff = 0.05 \nFC cutoff = 1"),
                pCutoff = 0.05,
                FCcutoff = 1,
                pointSize = 1.0,
                labSize = 4.0,
                col=c('black', 'black', 'black', 'red3'),
                ylim = c(0,y1),
                ylab = bquote(~-Log[10]~ 'FDR'),
                drawConnectors = TRUE,border = 'full',colConnectors = 'grey30',
                widthConnectors = .5, borderWidth = 1.0, borderColour = 'black') + theme(legend.position="none")

print(plot)

```


## Tg P301s tdTomato vs  All other 
```{r Tg Tdtomato vs all, message=FALSE, warning=FALSE, echo=TRUE}
# Read in Differential expression
data <- read_delim("TG_tdTomvsALL.csv", 
                       delim = ";", escape_double = FALSE,col_types = cols(log2FC = col_number()),
                       locale = locale(decimal_mark = ",", grouping_mark = "."),
                       trim_ws = TRUE)
```

```{r tgtdTomato vs all Volcano, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11.75, fig.height=10.04}
x <- data$GeneName
data <- data[-which(startsWith(as.character(x),"Krt")==TRUE),]
y1 <- ceiling(max(-log10(data$adj.pvalue))*2)/2
plot <- EnhancedVolcano(data,
                lab = data$GeneName,
                x = 'log2FC',
                y = 'adj.pvalue',
                title = 'TG_tdTom vs. All', caption='',
                subtitle = paste("FDR cutoff = 0.05 \nFC cutoff = 1"),
                pCutoff = 0.05,
                FCcutoff = 1,
                pointSize = 1.0,
                labSize = 4.0,
                col=c('black', 'black', 'black', 'red3'),
                ylim = c(0,y1),
                ylab = bquote(~-Log[10]~ 'FDR'),
                drawConnectors = TRUE,border = 'full',colConnectors = 'grey30',
                widthConnectors = .5, borderWidth = 1.0, borderColour = 'black') + theme(legend.position="none")

print(plot)

```

## Wt tdTomato vs All others

```{r WT tdT vs all, message=FALSE, warning=FALSE, echo=TRUE}
# Read in Differential expression
data <- read_delim("WtTdvsALL.csv", 
                       delim = ";", escape_double = FALSE,col_types = cols(log2FC = col_number()),
                       locale = locale(decimal_mark = ",", grouping_mark = "."),
                       trim_ws = TRUE)
```

```{r Wt tdT vs all Volcano, echo=FALSE, message=FALSE, warning=FALSE, fig.width=11.75, fig.height=10.04}
x <- data$GeneName
data <- data[-which(startsWith(as.character(x),"Krt")==TRUE),]
y1 <- ceiling(max(-log10(data$adj.pvalue))*2)/2
plot <- EnhancedVolcano(data,
                lab = data$GeneName,
                x = 'log2FC',
                y = 'adj.pvalue',
                title = 'WTtdTomato vs. All', caption='',
                subtitle = paste("FDR cutoff = 0.05 \nFC cutoff = 1"),
                pCutoff = 0.05,
                FCcutoff = 1,
                pointSize = 1.0,
                labSize = 4.0,
                col=c('black', 'black', 'black', 'red3'),
                ylim = c(0,y1),
                ylab = bquote(~-Log[10]~ 'FDR'),
                drawConnectors = TRUE,border = 'full',colConnectors = 'grey30',
                widthConnectors = .5, borderWidth = 1.0, borderColour = 'black') + theme(legend.position="none")

print(plot)

```